#DE_APA
```{r,eval=F,echo=F}


```{r filter res}




```

```{r pca}
library(data.table)
library(factoextra)
library(ggplot2)
library(FactoMineR)
sample.info <- fread('./Hema_sampleinfo.txt',header=T)
filtered <- fread('./DE_APA/filtered.RUD.txt',header = T)


tmp <- as.data.frame(filtered[,77:110])
tmp[tmp>1] <- 1
tmp[is.na(tmp)] <- 0
pca <- FactoMineR::PCA(t(tmp), ncp = 100, graph = F)
pca_result <- data.frame(pca$svd$U, ID = c('LT-HSC_1','LT-HSC_2',
                                           'ST-HSC_1','ST-HSC_2',
                                           gsub('.RUD','',colnames(tmp)[5:34])))
pca_result <- merge(pca_result, sample.info[, c("sample", "CellType", "Lineage")],
                    by.x = "ID",by.y='sample')

pcaplot<-ggplot(pca_result,aes(x = X1,y = X2,z=X3, col = CellType,pt.size=6))+
  geom_point(size=6,aes(shape=factor(Lineage)))+
  geom_text(aes(label=ID),hjust=0, vjust=0)+
  #theme_classic()+
  #guides(colour=FALSE)+
  xlab(paste("PC1(",round(pca$eig[,2][1],2),"%)",sep = ""))+
  ylab(paste("PC2(",round(pca$eig[,2][2],2),"%)",sep = "")) +
  labs(title = "RUD of Samples", subtitle = "Principal Component Analysis")+
  theme(plot.title = element_text(hjust = 0.5),
        #legend.position='bottom',
        plot.subtitle = element_text(hjust = 0.5))


  
pcaplot
```  
  
```{r 3d, eval=F,echo=FALSE}
library(gg3D)
ggplot(pca_result, aes(x=X1, y=X2, z=X3, color=CellType,shape=Lineage)) +
  axes_3D(phi=70) +
  stat_3D(phi=70,size=5) +
  #axis_labs_3D(phi=20)+
  labs_3D(phi=70,
          labs=c(paste("PC1(",round(pca$eig[,2][1],2),"%)",sep = ""), 
                 paste("PC2(",round(pca$eig[,2][2],2),"%)",sep = ""), 
                 paste("PC3(",round(pca$eig[,2][3],2),"%)",sep = "")),
           hjust=c(0,1,0.8), vjust=c(1, 1, -0.5), angle=c(30, -30, 90)) +
  theme_void()+
    labs(title = "RUD of Samples", subtitle = "Principal Component Analysis")+
  theme(plot.title = element_text(hjust = 0.5,size=14),
        #legend.position='bottom',
        plot.subtitle = element_text(hjust = 0.5,size=14))
ggsave(file='./Figures/pca.pdf',width = 8,height = 7)


fviz_screeplot(pca, addlabels = TRUE)
ggsave(file='./Figures/pca_contribution.pdf',width = 6,height = 6)

```



```{r randomforest}
library(varSelRF)
library(pheatmap)
genes.names <- do.call(c,lapply(strsplit(filtered$PA.ID,'\\|'),function(x)x[2]))
PA <- filtered$PA.ID
genes.id <- do.call(c,lapply(strsplit(filtered$PA.ID,'\\|'),function(x)x[1]))
x<-t(tmp)
cl<-as.factor(as.numeric(as.factor(sample.info$Lineage)))
set.seed(2)
selected<- varSelRF(x,cl,ntree=2000,ntreeIterat=1000,vars.drop.frac=0.2)
selected.var <- as.numeric(gsub('v','',selected$selected.vars))
selected.dt <- tmp[selected.var,]
rownames(selected.dt) <- PA[selected.var]
p <- pheatmap(selected.dt,main = 'Selected Variables using RandomForest',
              color = colorRampPalette(c("navy", "white", "firebrick3"))(200),
              fontsize_row=6)
pdf('Selected_Var.Lineage.pdf',height = 7,width = 12)
p
dev.off()

x<-t(tmp)
cl<-as.factor(as.numeric(as.factor(sample.info$CellType)))
set.seed(2)
selected1<- varSelRF(x,cl,ntree=2000,ntreeIterat=1000,vars.drop.frac=0.2)
selected1.var <- as.numeric(gsub('v','',selected1$selected.vars))
selected1.dt <- tmp[selected1.var,]
rownames(selected1.dt) <- PA[selected1.var]
p <- pheatmap(selected1.dt,main = 'Selected Variables using RandomForest',
              color = colorRampPalette(c("navy", "white", "firebrick3"))(200),
              fontsize_row=6)

pdf('Selected_Var.pdf',height = 4,width = 6)
p
dev.off()
library(RColorBrewer)
#gene expression
GeneExp <- fread('/mnt/raid63/HSC/mouse/bulk/chenghui_MmBlood/meta_RD/Mm_chenghui_hemo.RSEM_Gene_TPM.tsv',header = T)

Gene.use <- GeneExp[gene_id%in%genes.id[selected.var],]
Gene.use <- as.data.frame(Gene.use[,-1])
rownames(Gene.use) <- genes.names[selected.var]

library(circlize)
library(grid)
library(ComplexHeatmap)
ht1 = Heatmap(selected.dt, name = "RUD",column_title = "RUD of samples",
              col=colorRampPalette(c("navy", "white", "firebrick3"))(150))
# 批量按行中心标准化，减均值除方差，Z-score,因为我们关注基因在样本间的变化，而不是样本内的基因的变化

ht2 = Heatmap(t(scale(t(Gene.use))),mat_scaled, name = "Normalized TPM", 
              column_title = "Gene Expression", 
              col=colorRampPalette(c("navy", "white", "firebrick3"))(150))

 pheatmap(Gene.use, main="Gene Expression",
          color = colorRampPalette(c("navy", "white", "firebrick3"))(200),
         # color  = colorRampPalette(brewer.pal(3, "Set1"))(256),
         scale = 'row')
 
 pheatmap(t(scale(t(Gene.use))), main="Gene Expression",
          color = colors,
          breaks = breaks)
 
 generate_breaks = function(x, n, center = F){
    if(center){
        m = max(abs(c(min(x, na.rm = T), max(x, na.rm = T))))
        res = seq(-m, m, length.out = n + 1)
    }
    else{
        res = seq(min(x, na.rm = T), max(x, na.rm = T), length.out = n + 1)
    }
    
    return(res)
}
 
 colors = colorRampPalette(c("navy", "white", "firebrick3"))(200)
 breaks = generate_breaks(t(scale(t(Gene.use))),length(colors),center = T)

#region for sashimiplot
region <- filtered[selected.var,]
region <- paste0(region$chr,':',region$start,':',
                 region$end,'.',genes.names[selected.var],'.',region$PA.x)

write.table(region,file='./SashimiPlot/region',col.names=F,row.names=F,quote=F)

GO_genes <- unique(genes.id[selected.var])
library(AnnotationDbi)
library(org.Mm.eg.db)
library(DOSE)
library(GO.db)
library(topGO)
library(GSEABase)
library(clusterProfiler)
library(Rgraphviz)
PA.GO <- enrichGO(gene = GO_genes,
                       OrgDb  = org.Mm.eg.db, keyType = "ENSEMBL",
                       ont  = "BP", pAdjustMethod = "BH",
                       pvalueCutoff  = 0.05,qvalueCutoff  = 0.05)
symbols <- unique(genes.names[selected.var])
KEGG_genes<-mapIds(org.Mm.eg.db, symbols, 'ENTREZID', 'SYMBOL')
PA.KEGG <- enrichKEGG(KEGG_genes,organism='mmu')
pdf(paste('GO_lineage.pdf'),height = 8,width = 10)
dotplot(PA.GO,showCategory = 20,font.size = 10)
enrichMap(PA.GO)
plotGOgraph(PA.GO,firstSigNodes = 15, useInfo = "all")
dev.off()


region1 <- filtered[selected1.var,]
region1 <- paste0(region1$chr,':',region1$start,':',
                 region1$end,'.',genes.names[selected1.var],'.',region1$PA.x)

write.table(region1,file='./SashimiPlot/region_cell',col.names=F,row.names=F,quote=F)
```



```{r log2FC deltaRUD}
#all cell type compared with LT-HSC,test on all genes before filtering
cells<- sample.info$CellType
cell.mtx<- list()
for(i in cells){
  samples <- sample.info[CellType==i,]$sample
  cell.columns <- c(paste0(samples,'.L_Coverage'),paste0(samples,'.S_Coverage'),
                   paste0(samples,'.RUD'))
  mtx <- newfiles[,..cell.columns] 
  mtx[is.na(mtx)] <- 0
  mtx$mean_L <- rowMeans(mtx[,1:2])
  mtx$mean_S <- rowMeans(mtx[,3:4])
  mtx$mean_RUD <- rowMeans(mtx[,5:6])
  cell.mtx[[i]] <- mtx
}


#evaluate the statistical significance of RUD difference
Test.St <- function(x,y){
  t.mtx <- cbind(x[,7:8],y[,7:8])
  suppressWarnings(p_value <- apply(t.mtx,1,function(x){
    con.mtx <- matrix(x,nrow=2)
    p.v <- fisher.test(con.mtx)$p.value
    return(p.v)
  }))
  FDR <- p.adjust(p_value)
  delta_RUD <- x[,9] - y[,9]
  log2FC <- abs(log2(x[,9]/y[,9]))
  RUD_info <- rowMeans(cbind(x[,9],y[,9]),na.rm = T)
  res <- cbind(p_value,FDR,delta_RUD,log2FC,RUD_info)
  colnames(res) <- c('P_value','FDR','Delta_RUD','Log2FC','RUD_info')
  return(res)
}

test.1 <- Test.St(x=cell.mtx[[1]],y=cell.mtx[[2]])
test.2 <- Test.St(x=cell.mtx[[1]],y=cell.mtx[[3]])
test.3 <- Test.St(x=cell.mtx[[1]],y=cell.mtx[[4]])
test.4 <- Test.St(x=cell.mtx[[1]],y=cell.mtx[[5]])
test.5 <- Test.St(x=cell.mtx[[1]],y=cell.mtx[[6]])
test.6 <- Test.St(x=cell.mtx[[1]],y=cell.mtx[[7]])
test.7 <- Test.St(x=cell.mtx[[1]],y=cell.mtx[[8]])
test.8 <- Test.St(x=cell.mtx[[1]],y=cell.mtx[[9]])
test.9 <- Test.St(x=cell.mtx[[1]],y=cell.mtx[[10]])
test.10 <- Test.St(x=cell.mtx[[1]],y=cell.mtx[[11]])
test.11 <- Test.St(x=cell.mtx[[1]],y=cell.mtx[[12]])
test.12 <- Test.St(x=cell.mtx[[1]],y=cell.mtx[[13]])
test.13 <- Test.St(x=cell.mtx[[1]],y=cell.mtx[[14]])
test.14 <- Test.St(x=cell.mtx[[1]],y=cell.mtx[[15]])
test.15 <- Test.St(x=cell.mtx[[1]],y=cell.mtx[[16]])
test.16 <- Test.St(x=cell.mtx[[1]],y=cell.mtx[[17]])



use.1 <- test.1[FDR <= 0.05&Delta_RUD >= 0.1&Log2FC >= 0.5,]

dt <- newfiles[PA.ID%in%use$ID,c(69,70,73,74)]
dt[is.na(dt)] <- 0
dt[dt>1] <- 1
dt <- dt[rowMeans(dt) <1,]

pheatmap(dt,cluster_cols = F)
```


